Ordinal Models

### 2024-12-08 docs/misc/log_error.qmd
big update to results section. Added a few paragraphs to Intro.

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

@@

>>>>

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

@@

# S2

log_error_model_s2_c1 <- brm(
  log_abs_error ~ refClass*calc + (1|id) + (1|state),
  data = s2_agg1,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 2000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1.rds')) 
)

summary(log_error_model_s2_c1)
conditional_effects(log_error_model_s2_c1)

log_error_model_s2_c1rf <- brm(
  log_abs_error ~ refClass + (1|id) + (1|state) + (1|calc),
  data = s2_agg1,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 2000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf.rds')) 
)

summary(log_error_model_s2_c1rf)

log_error_model_s2_cAdd_int3 <- brm(
  log_abs_error ~ (refClass*rounded*pct_goal)+calc + (1|id) + (1|state),
  data = s2_agg1,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s2_cAdd_int3.rds')) 
)

summary(log_error_model_s2_cAdd_int3)
conditional_effects(log_error_model_s2_cAdd_int3)

log_error_model_s2_c1rf_int3 <- brm(
  log_abs_error ~ refClass*rounded*pct_goal + (1|id) + (1|state) + (1|calc),
  data = s2_agg1,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf_int3.rds')) 
)

summary(log_error_model_s2_c1rf_int3)

conditional_effects(log_error_model_s2_c1rf_int3)

log_error_model_s2_c1rf_add3 <- brm(
  log_abs_error ~ refClass+rounded+pct_goal + (1|id) + (1|state) + (1|calc),
  data = s2_agg1,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf_add3.rds')) 
)

summary(log_error_model_s2_c1rf_add3)
conditional_effects(log_error_model_s2_c1rf_add3)
